Extracting The Power of Dependent Types

### 2021-07-13 paper.tex
typos

In this paper we investigate the idea of developing
embedded programs hand-in-hand with custom
code generators for them. We solve the well-known
conundrum of choosing between deep and shallow
embedding by leveraging the power of
reflection.
This allows us to enjoy simplicity of writing progams
in shallow embedding.  Yet we keep full access to the internal
structure of the embedded programs, so that we could translate
them into the original language.

We apply this idea in the context of dependently-typed
language Agda.  We demonstrate the use of dependent types
in order to enforce static propeties of the embedded programs.
Concretely, we have demonstrated the approach by implementing
an extractor for a fragment of the PostScript language.

The main advantage of our approach is twofold: first, it allows
you to implement all these things --- as well as the embedded programs
themselves --- \emph{side by side in a single language}, simplifying
the development. And secondly, it allows you to reuse the typechecker
and evaluator of the host language, saving you a lot of work.
While we showcase a single example, we applied our approach to several other
languages:
\begin{itemize}
\item Kaleidoscope, a minimalist imperative language used in the
  documentation of LLVM.
\item SaC (Single assignment C), a high-performance array language
  that generates efficient code for various architectures.
\item APL (Array Programming Language), another array language with
  heavily overloaded syntax that pioneered the concept of rank- and
  shape-polymorphic programming.
\end{itemize}
For more details on these examples,
refer to~\citet{DBLP:journals/corr/abs-2105-10819}.
This sugests that you are now empowered to apply the proposed
technique tp verify code in any language.

>>>>

In this paper we investigate the idea of developing
embedded programs hand-in-hand with custom
code generators for them. We solve the well-known
conundrum of choosing between deep and shallow
embedding by leveraging the power of
reflection.
This allows us to enjoy simplicity of writing programs
in shallow embedding.  Yet we keep full access to the internal
structure of the embedded programs, so that we could translate
them into the original language.

We apply this idea in the context of dependently-typed
language Agda.  We demonstrate the use of dependent types
in order to enforce static properties of the embedded programs.
Concretely, we have demonstrated the approach by implementing
an extractor for a fragment of the PostScript language.

The main advantage of our approach is twofold: first, it allows
you to implement all these things --- as well as the embedded programs
themselves --- \emph{side by side in a single language}, simplifying
the development. And secondly, it allows you to reuse the typechecker
and evaluator of the host language, saving you a lot of work.
While we showcase a single example, we applied our approach to several other
languages:
\begin{itemize}
\item Kaleidoscope, a minimalist imperative language used in the
  documentation of LLVM.
\item SaC (Single assignment C), a high-performance array language
  that generates efficient code for various architectures.
\item APL (Array Programming Language), another array language with
  heavily overloaded syntax that pioneered the concept of rank- and
  shape-polymorphic programming.
\end{itemize}
For more details on these examples,
refer to~\citet{DBLP:journals/corr/abs-2105-10819}.
This suggests that you are now empowered to apply the proposed
technique to verify code in any language.
